### 2022-02-19 misc/solutions/prob-method-sols.tex
3.3

>>>>

\proclaim Exercise 3.3. Prove that every $3$-uniform hypergraph with $n$ vertices and $m\ge n/3$ edges
contains an independent set (i.e., a set of vertices containing no edges) of size at least
$${2n^{3/2}\over 3\sqrt 3\sqrt m}.$$

\proof We mirror the proof of Theorem~3.2.1. Let $(X\cap \F)$ be a $3$-uniform hypergraph with
$|X| = n$ and $|\F| = m \ge n/3$. Let $S\subseteq X$ be a random subset obtained by independently
putting each $x\in X$ in $S$ with probability $p$, to be chosen later. Of course, $\ex\bigl\{|X|\bigr\} = np$.
What is the expected number of edges contained in $S$? Well, each edge $\{x,y,z\}\in \F$ has probability
$p^3$ of being contained in $S$, so we expect $mp^3$ edges in the subgraph induced by $S$. If
$Y$ is the set of these edges, we now want to choose $p$ to
minimise the quantity $\ex\bigl\{|X|-|Y|\bigr\} = np-mp^3$. The derivative
is $n-3mp^2$ so we ought to set $p=\sqrt{n/3m}$, obtaining
$$\ex\bigl\{ |X|-|Y|\bigr\} = {n^{3/2}-m\sqrt n\over \sqrt{3m}}.$$
Since $m\ge n/3$, we have
$$\ex\bigl\{ |X|-|Y|\bigr\} \ge {n^{3/2}\over \sqrt{3m}}-{n^{3/2}\over 3\sqrt{3m}}={2n^{3/2}\over 3\sqrt{3m}}.$$
This means there exists some set $S\subseteq X$ with the number of vertices minus the number of edges
at least $2n^{3/2}/\bigl(3\sqrt{3m}\bigr)$. Removing one vertex from each edge, we arrive at an independent
set of vertices $S'$ with cardinality at least this quantity.\slug
